A novel distributed nature-inspired algorithm for solving optimization problems\thanks{Supported by TECNOLÓGICO NACIONAL DE MÉXICO (TecNM).}

### 2021-09-28 paper.tex
Work in progress
Experiment tables update

This algorithm seeks to imitate the natural life cycle, where new individuals
are born at any moment and mature over time, where they age and suffer
mutations throughout their lives. In reproduction, couples match by mutual
attraction, where they may have offspring. Death can happen to everyone: from a
newborn to an aged adult, where the individual's fitness will impact their
longevity.

@@

The figure 4 describes the general flowchart for the Life-Cycle algorithm.

@@

One strategic difference in our algorithm implementation is the reproduction
process, which is flexible to work in parallel with multiple couple selection
methods, for example, tournament selection, random selection (closed
reproduction), and couple match by creating a new mating individual (open
reproduction). We use many combinations of the containerized processes, up to
the minimum number required to obtain good results. Figure 5 shows a comparison
of only two alternatives for the implementation of the algorithm. Our
experimentation started with a configuration of ten processes and ended in
five, shown on the left and right sides of figure 5, respectively.

@@

Experiment 1: The goal of our first OneMax experiment was to make sure our
Life-Cycle algorithm worked as expected, converging on the solution. For this
experiment, we intentionally used delays (on the tenth of a second magnitude)
to follow the Life-Cycle algorithm behavior on the console output. For this
experiment, we compared the DEAP algorithm versus the Life-Cycle using ten
processes (on containers) where we show the summarized results in Table 2, and
Figure 6 shows their Box and Whisker chart.

@@

Experiment 2: The goal of our second experiment was to confirm the Life-Cycle
algorithm continued working as expected, converging on the solution. For this
experiment, we reduced the time used on delays (now on the thousands of a
second magnitude) to follow the Life-Cycle algorithm behavior. For this
experiment, we only used tournament selection for the reproduction, on the
first configuration running the Life-Cycle on ten processes, versus the second
configuration where we reduced the processes to the minimum basic 5. We show
the summarized results in Table 3, and Figure 7 shows their Box and Whisker
chart.

@@

Experiment 3: The goal of our third experiment was to follow and study the
behavior of the Life-Cycle algorithm that continued converging on the solution.
For this experiment, we remained using minimal delays (thousands of a second
magnitude). For this experiment, we only used tournament selection for the
reproduction, on the first configuration running the Life-Cycle on six
processes, two of whom was the Death stage, versus the second configuration
where we reduced the processes to the minimum (five) but increasing the (Death)
goal approval range to 115. We show the summarized results in Table 4, and
Figure 8 shows their Box and Whisker chart.

@@

Experiment 4: The goal of our fourth and last experiment was to follow and
study the behavior of the Life-Cycle algorithm that continued converging on the
solution. For this experiment, we remained using minimal delays (thousands of a
second magnitude). For this experiment, we compared the OneMax DEAP
implementation versus the Life-Cycle algorithm, only using tournament selection
for the reproduction, with the Life-Cycle configuration running on the minimum
(five) processes but increasing the (Death) goal approval range to 125. We show
the summarized results in Table 5, and Figure 9 shows their Box and Whisker
chart.

@@

    \begin{table}[]
        \centering        
        \caption{OneMax DEAP experiment results.}\label{tab3}        
        \begin{tabular}{|l|l|l|l|l|}
        \hline
        \multicolumn{5}{|c|}{\textbf{OneMax DEAP}} \\ \hline
        \begin{tabular}[c]{@{}l@{}}Run   \\    \\ No.\end{tabular} & Generation & \begin{tabular}[c]{@{}l@{}}Best \\    \\ Individual\end{tabular} & \begin{tabular}[c]{@{}l@{}}Total\\    \\ Evaluations\end{tabular} & \begin{tabular}[c]{@{}l@{}}Total Time \\    \\ (seconds)\end{tabular} \\ \hline
        1 & 6 & 20 & 360 & 0.016 \\ \hline
        2 & 5 & 20 & 300 & 0.032 \\ \hline
        3 & 7 & 20 & 420 & 0.061 \\ \hline
        4 & 9 & 20 & 540 & 0.081 \\ \hline
        5 & 8 & 20 & 480 & 0.080 \\ \hline
        6 & 6 & 20 & 360 & 0.062 \\ \hline
        7 & 7 & 20 & 420 & 0.062 \\ \hline
        8 & 6 & 20 & 360 & 0.016 \\ \hline
        9 & 8 & 20 & 480 & 0.058 \\ \hline
        10 & 6 & 20 & 360 & 0.050 \\ \hline
        11 & 7 & 20 & 420 & 0.073 \\ \hline
        12 & 8 & 20 & 480 & 0.076 \\ \hline
        13 & 8 & 20 & 480 & 0.065 \\ \hline
        14 & 6 & 20 & 360 & 0.060 \\ \hline
        15 & 6 & 20 & 360 & 0.045 \\ \hline
        16 & 6 & 20 & 360 & 0.059 \\ \hline
        17 & 6 & 20 & 360 & 0.061 \\ \hline
        18 & 6 & 20 & 360 & 0.058 \\ \hline
        19 & 5 & 20 & 300 & 0.060 \\ \hline
        20 & 7 & 20 & 420 & 0.067 \\ \hline
        21 & 7 & 20 & 420 & 0.060 \\ \hline
        22 & 7 & 20 & 420 & 0.085 \\ \hline
        23 & 7 & 20 & 420 & 0.071 \\ \hline
        24 & 4 & 20 & 300 & 0.049 \\ \hline
        25 & 6 & 20 & 360 & 0.050 \\ \hline
        26 & 9 & 20 & 540 & 0.081 \\ \hline
        27 & 6 & 20 & 420 & 0.074 \\ \hline
        28 & 6 & 20 & 360 & 0.064 \\ \hline
        29 & 6 & 20 & 360 & 0.065 \\ \hline
        30 & 4 & 20 & 240 & 0.016 \\ \hline
        Avg. & 6.5 & 20 & 394 & 0.059 \\ \hline
        \end{tabular}
        \end{table}

    \begin{table}[]
        \centering        
        \caption{LifeCycle Algorithm experiment results.}\label{tab4}
        \begin{tabular}{|l|l|l|l|l|}
        \hline
        \multicolumn{5}{|c|}{\textbf{LifeCycle Algorithm}} \\ \hline
        \begin{tabular}[c]{@{}l@{}}Run   \\    \\ No.\end{tabular} & Generation & \begin{tabular}[c]{@{}l@{}}Best \\    \\ Individual\end{tabular} & \begin{tabular}[c]{@{}l@{}}Total\\    \\ Evaluations\end{tabular} & \begin{tabular}[c]{@{}l@{}}Total Time \\    \\ (seconds)\end{tabular} \\ \hline
        1 & 8 & 20 & 486 & 6.65 \\ \hline
        2 & 6 & 20 & 381 & 5.20 \\ \hline
        3 & 6 & 20 & 369 & 5.69 \\ \hline
        4 & 7 & 20 & 434 & 6.37 \\ \hline
        5 & 12 & 20 & 731 & 8.20 \\ \hline
        6 & 5 & 20 & 306 & 4.55 \\ \hline
        7 & 13 & 20 & 764 & 8.34 \\ \hline
        8 & 4 & 20 & 233 & 4.09 \\ \hline
        9 & 12 & 20 & 692 & 7.27 \\ \hline
        10 & 6 & 20 & 338 & 4.76 \\ \hline
        11 & 9 & 20 & 559 & 7.78 \\ \hline
        12 & 12 & 20 & 730 & 8.35 \\ \hline
        13 & 4 & 20 & 219 & 3.43 \\ \hline
        14 & 8 & 20 & 462 & 5.28 \\ \hline
        15 & 4 & 20 & 217 & 3.37 \\ \hline
        16 & 4 & 20 & 234 & 4.54 \\ \hline
        17 & 4 & 20 & 256 & 4.35 \\ \hline
        18 & 12 & 20 & 714 & 8.14 \\ \hline
        19 & 12 & 20 & 694 & 7.88 \\ \hline
        20 & 15 & 20 & 892 & 9.74 \\ \hline
        21 & 11 & 20 & 633 & 7.98 \\ \hline
        22 & 3 & 20 & 169 & 4.02 \\ \hline
        23 & 10 & 20 & 611 & 6.89 \\ \hline
        24 & 10 & 20 & 629 & 7.32 \\ \hline
        25 & 7 & 20 & 416 & 6.22 \\ \hline
        26 & 7 & 20 & 446 & 5.55 \\ \hline
        27 & 6 & 20 & 379 & 5.95 \\ \hline
        28 & 8 & 20 & 466 & 6.98 \\ \hline
        29 & 3 & 20 & 197 & 3.36 \\ \hline
        30 & 6 & 20 & 381 & 6.16 \\ \hline
        Avg. & 7.8 & 20 & 468 & 6.15 \\ \hline
        \end{tabular}
        \end{table}

>>>>

This algorithm seeks to imitate the natural life cycle, where new individuals
are born at any moment and mature over time, where they age and suffer
mutations throughout their lives. In reproduction, couples match by mutual
attraction, where they may have offspring. Death can happen to everyone: from a
newborn to an aged adult, where the individual's fitness will impact their
longevity. The general model concept is shown on the Fig.~\ref{fig2}.

@@

The Figure~\ref{fig4} describes the general flowchart for the Life-Cycle algorithm.

@@

One strategic difference in our algorithm implementation is the reproduction
process, which is flexible to work in parallel with multiple couple selection
methods, for example, tournament selection, random selection (closed
reproduction), and couple match by creating a new mating individual (open
reproduction). We use many combinations of the containerized processes, up to
the minimum number required to obtain good results. Figure~\ref{fig5} shows a comparison
of only two alternatives for the implementation of the algorithm. Our
experimentation started with a configuration of ten processes and ended in
five, shown on the left and right sides of Fig.~\ref{fig5}, respectively.

@@

Experiment 1: The goal of our first OneMax experiment was to make sure our
Life-Cycle algorithm worked as expected, converging on the solution. For this
experiment, we intentionally used delays (on the tenth of a second magnitude)
to follow the Life-Cycle algorithm behavior on the console output. For this
experiment, we compared the DEAP algorithm versus the Life-Cycle using ten
processes (on containers) where we show the summarized results in Table~\ref{tab2}, and
Figure~\ref{fig6} shows their Box and Whisker chart.

\begin{table}[]
    \centering        
    \caption{Experiment 1 results: OneMax DEAP versus Life-Cycle (10 processes).}\label{tab2}
    \begin{tabular}{|l|l|l|l|l|l|l|l|l|}
    \hline
    Run & \multicolumn{4}{l|}{OneMax DEAP} & \multicolumn{4}{l|}{Life-Cycle (10P)} \\ \hline
    1-60 & Last Gen. & Eval. & Time (sec) & Eval/sec & Last Gen. & Eval. & Time (sec) & Eval/sec \\ \hline
    Avg. & 6.5 & 391 & 0.036 & 10,836 & 9.6 & 578 & 6.830 & 85 \\ \hline
    \end{tabular}
    \end{table}

@@

Experiment 2: The goal of our second experiment was to confirm the Life-Cycle
algorithm continued working as expected, converging on the solution. For this
experiment, we reduced the time used on delays (now on the thousands of a
second magnitude) to follow the Life-Cycle algorithm behavior. For this
experiment, we only used tournament selection for the reproduction, on the
first configuration running the Life-Cycle on ten processes, versus the second
configuration where we reduced the processes to the minimum basic 5. We show
the summarized results in Table~\ref{tab3}, and Figure~\ref{fig7} shows their Box and Whisker
chart.

\begin{table}[]
    \centering        
    \caption{Experiment 2 results, Life-Cycle Tournament selection: 10 versus 5 processes.}\label{tab3}
    \begin{tabular}{|l|l|l|l|l|l|l|l|l|}
    \hline
    Run & \multicolumn{4}{l|}{LifeCycle (Tourn. 10P)} & \multicolumn{4}{l|}{LifeCycle (Tourn. 5P)} \\ \hline
    1-60 & Last Gen. & Eval. & Time (sec) & Eval/sec & Last Gen. & Eval. & Time (sec) & Eval/sec \\ \hline
    Avg. & 7.2 & 434 & 0.740 & 587 & 8.2 & 495 & 0.826 & 599 \\ \hline
    \end{tabular}
    \end{table}

@@

Experiment 3: The goal of our third experiment was to follow and study the
behavior of the Life-Cycle algorithm that continued converging on the solution.
For this experiment, we remained using minimal delays (thousands of a second
magnitude). For this experiment, we only used tournament selection for the
reproduction, on the first configuration running the Life-Cycle on six
processes, two of whom was the Death stage, versus the second configuration
where we reduced the processes to the minimum (five) but increasing the (Death)
goal approval range to 115. We show the summarized results in Table~\ref{tab4}, and
Figure~\ref{fig8} shows their Box and Whisker chart.

\begin{table}[]
    \centering        
    \caption{Experiment 3 results, Life-Cycle Tournament: 6 processes (Death x2) versus 5 processes (80 to 115 goal).}\label{tab4}
    \begin{tabular}{|l|l|l|l|l|l|l|l|l|}
    \hline
    Run & \multicolumn{4}{l|}{LifeCycle (Tourn. 6P, Death x2)} & \multicolumn{4}{l|}{LifeCycle (Tourn. 5P, 80-115g)} \\ \hline
    1-60 & Last Gen. & Eval. & Time (sec) & Eval/sec & Last Gen. & Eval. & Time (sec) & Eval/sec \\ \hline
    Avg. & 8.0 & 483 & 0.846 & 571 & 6.4 & 384 & 0.695 & 553 \\ \hline
    \end{tabular}
    \end{table}

@@

Experiment 4: The goal of our fourth and last experiment was to follow and
study the behavior of the Life-Cycle algorithm that continued converging on the
solution. For this experiment, we remained using minimal delays (thousands of a
second magnitude). For this experiment, we compared the OneMax DEAP
implementation versus the Life-Cycle algorithm, only using tournament selection
for the reproduction, with the Life-Cycle configuration running on the minimum
(five) processes but increasing the (Death) goal approval range to 125. We show
the summarized results in Table~\ref{tab5}, and Figure~\ref{fig9} shows their Box and Whisker
chart.

\begin{table}[]
    \centering        
    \caption{Experiment 4 results, OneMax DEAP versus Life-Cycle Tournament: 5 processes (80 to 125 goal).}\label{tab5}
    \begin{tabular}{|l|l|l|l|l|l|l|l|l|}
    \hline
    Run & \multicolumn{4}{l|}{OneMax DEAP} & \multicolumn{4}{l|}{Life-Cycle (Tourn. 5P, 80-125g)} \\ \hline
    1-60 & Last Gen. & Eval. & Time (sec) & Eval/sec & Last Gen. & Eval. & Time (sec) & Eval/sec \\ \hline
    Avg. & 6.5 & 391 & 0.036 & 10,836 & 6.3 & 377 & 0.678 & 556 \\ \hline
    \end{tabular}
    \end{table}

@@

    \begin{table}[]
        \centering        
        \caption{OneMax DEAP experiment results.}\label{tab6}        
        \begin{tabular}{|l|l|l|l|l|}
        \hline
        \multicolumn{5}{|c|}{\textbf{OneMax DEAP}} \\ \hline
        \begin{tabular}[c]{@{}l@{}}Run   \\    \\ No.\end{tabular} & Generation & \begin{tabular}[c]{@{}l@{}}Best \\    \\ Individual\end{tabular} & \begin{tabular}[c]{@{}l@{}}Total\\    \\ Evaluations\end{tabular} & \begin{tabular}[c]{@{}l@{}}Total Time \\    \\ (seconds)\end{tabular} \\ \hline
        1 & 6 & 20 & 360 & 0.016 \\ \hline
        2 & 5 & 20 & 300 & 0.032 \\ \hline
        3 & 7 & 20 & 420 & 0.061 \\ \hline
        4 & 9 & 20 & 540 & 0.081 \\ \hline
        5 & 8 & 20 & 480 & 0.080 \\ \hline
        6 & 6 & 20 & 360 & 0.062 \\ \hline
        7 & 7 & 20 & 420 & 0.062 \\ \hline
        8 & 6 & 20 & 360 & 0.016 \\ \hline
        9 & 8 & 20 & 480 & 0.058 \\ \hline
        10 & 6 & 20 & 360 & 0.050 \\ \hline
        11 & 7 & 20 & 420 & 0.073 \\ \hline
        12 & 8 & 20 & 480 & 0.076 \\ \hline
        13 & 8 & 20 & 480 & 0.065 \\ \hline
        14 & 6 & 20 & 360 & 0.060 \\ \hline
        15 & 6 & 20 & 360 & 0.045 \\ \hline
        16 & 6 & 20 & 360 & 0.059 \\ \hline
        17 & 6 & 20 & 360 & 0.061 \\ \hline
        18 & 6 & 20 & 360 & 0.058 \\ \hline
        19 & 5 & 20 & 300 & 0.060 \\ \hline
        20 & 7 & 20 & 420 & 0.067 \\ \hline
        21 & 7 & 20 & 420 & 0.060 \\ \hline
        22 & 7 & 20 & 420 & 0.085 \\ \hline
        23 & 7 & 20 & 420 & 0.071 \\ \hline
        24 & 4 & 20 & 300 & 0.049 \\ \hline
        25 & 6 & 20 & 360 & 0.050 \\ \hline
        26 & 9 & 20 & 540 & 0.081 \\ \hline
        27 & 6 & 20 & 420 & 0.074 \\ \hline
        28 & 6 & 20 & 360 & 0.064 \\ \hline
        29 & 6 & 20 & 360 & 0.065 \\ \hline
        30 & 4 & 20 & 240 & 0.016 \\ \hline
        Avg. & 6.5 & 20 & 394 & 0.059 \\ \hline
        \end{tabular}
        \end{table}

    \begin{table}[]
        \centering        
        \caption{LifeCycle Algorithm experiment results.}\label{tab7}
        \begin{tabular}{|l|l|l|l|l|}
        \hline
        \multicolumn{5}{|c|}{\textbf{LifeCycle Algorithm}} \\ \hline
        \begin{tabular}[c]{@{}l@{}}Run   \\    \\ No.\end{tabular} & Generation & \begin{tabular}[c]{@{}l@{}}Best \\    \\ Individual\end{tabular} & \begin{tabular}[c]{@{}l@{}}Total\\    \\ Evaluations\end{tabular} & \begin{tabular}[c]{@{}l@{}}Total Time \\    \\ (seconds)\end{tabular} \\ \hline
        1 & 8 & 20 & 486 & 6.65 \\ \hline
        2 & 6 & 20 & 381 & 5.20 \\ \hline
        3 & 6 & 20 & 369 & 5.69 \\ \hline
        4 & 7 & 20 & 434 & 6.37 \\ \hline
        5 & 12 & 20 & 731 & 8.20 \\ \hline
        6 & 5 & 20 & 306 & 4.55 \\ \hline
        7 & 13 & 20 & 764 & 8.34 \\ \hline
        8 & 4 & 20 & 233 & 4.09 \\ \hline
        9 & 12 & 20 & 692 & 7.27 \\ \hline
        10 & 6 & 20 & 338 & 4.76 \\ \hline
        11 & 9 & 20 & 559 & 7.78 \\ \hline
        12 & 12 & 20 & 730 & 8.35 \\ \hline
        13 & 4 & 20 & 219 & 3.43 \\ \hline
        14 & 8 & 20 & 462 & 5.28 \\ \hline
        15 & 4 & 20 & 217 & 3.37 \\ \hline
        16 & 4 & 20 & 234 & 4.54 \\ \hline
        17 & 4 & 20 & 256 & 4.35 \\ \hline
        18 & 12 & 20 & 714 & 8.14 \\ \hline
        19 & 12 & 20 & 694 & 7.88 \\ \hline
        20 & 15 & 20 & 892 & 9.74 \\ \hline
        21 & 11 & 20 & 633 & 7.98 \\ \hline
        22 & 3 & 20 & 169 & 4.02 \\ \hline
        23 & 10 & 20 & 611 & 6.89 \\ \hline
        24 & 10 & 20 & 629 & 7.32 \\ \hline
        25 & 7 & 20 & 416 & 6.22 \\ \hline
        26 & 7 & 20 & 446 & 5.55 \\ \hline
        27 & 6 & 20 & 379 & 5.95 \\ \hline
        28 & 8 & 20 & 466 & 6.98 \\ \hline
        29 & 3 & 20 & 197 & 3.36 \\ \hline
        30 & 6 & 20 & 381 & 6.16 \\ \hline
        Avg. & 7.8 & 20 & 468 & 6.15 \\ \hline
        \end{tabular}
        \end{table}
